---
title: "Course Project 1"
author: "Y. Xu"
date: "6/25/2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
data <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?
```{r}
total <- group_by(data, date) %>% summarise(total = sum(steps))
total_mean <- mean(total$total, na.rm = TRUE)
total_median <- median(total$total, na.rm = TRUE)
hist(total$total, main = 'Total Number of Steps Taken Each Day', xlab = 'Number of Steps', ylim = c(0,30))
```

The mean total number of steps taken per day is `r format(total_mean,round=2)`.
The median total number of steps taken per day is `r format(total_median,round=2)`.

## What is the average daily activity pattern?
```{r}
average <- group_by(data, interval) %>% summarise(average = mean(steps, na.rm = TRUE))
index <- which.max(average$average)
t <- average[index,1]
plot(average$interval, average$average, type = 'l', xlab = 'Interval', ylab = 'Number of Steps', main = 'Average Daily Activity Pattern')
```

Interval `r t` contains the maximum number of steps.

## Imputing missing values
```{r}
n <- sum(is.na(data$steps))
data1 <- data
for (i in 1:nrow(data1)){
    if (is.na(data1[i,1])) {
        idx <- which(data1[i,3]==average$interval)
        data1[i,1] <- average[idx,2]
    }
}
total1 <- group_by(data1, date) %>% summarise(total1 = sum(steps))
total1_mean <- mean(total1$total1)
total1_median <- median(total1$total1)
hist(total1$total1, main = 'Total Number of Steps Taken Each Day (Imputing Missing Values)', xlab = 'Number of Steps')
```

The total number of missing values in the dataset is `r n`.
The mean total number of steps taken per day is `r format(total1_mean,round=2)`.
The median total number of steps taken per day is `r format(total1_median,round=2)`.
These values do not differ from the estimates from the first part of the assignment.
Imputing missing data increases the frequency of the 10000-15000 steps, which is the group that the mean value belongs to. 


## Are there differences in activity patterns between weekdays and weekends?
```{r}
data1$week <- weekdays(as.Date(data1$date))
for (i in 1:nrow(data1)){
    if (data1[i,4]=='Saturday' | data1[i,4]=='Sunday') data1[i,4] <- 'weekend'
    else data1[i,4] <- 'weekday'
}
average1 <- group_by(data1, week, interval) %>% summarise(average1 = mean(steps))
qplot(interval, average1, data = average1, facets = week ~ ., geom = 'line', ylab = 'number of steps', main = 'Average Daily Activity Pattern - Weekday vs. Weekend')
```

